---
title: "Lube_city"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Creating seasonally cyclical models to estimate future monthly revenue. 
Please read the readme.file in this repository for more details on the goals of this project.



Load required packages:
```{r}

suppressMessages(library(ggfortify))
suppressMessages(library(ggplot2))
suppressMessages(library(fpp2))
suppressMessages(library(mgcv))
suppressMessages(library(forcats))
suppressMessages(library(dplyr))
suppressMessages(library(ggpubr))

```


Load data:
```{r}

data <- read.csv("Data.csv", stringsAsFactors = FALSE)

data2 <- data[!(data$Month == "2021_Nov"), ]


# Index the data into different data frames (training sets and total sets). We want to train the a seasonally cyclical model(s) on the monthly sales revenue from Oct 2020 - Sept 2021. Using these models we can then forecast revenue estimates for the period of interest (more info in the readme file). 

# Store 16 train model data:
data_16 <- data[!(data$Store_Id == "1" | data$Store_Id == "23" | 
                    data$Month == "2021_Nov" | data$Month == "2021_Oct" | 
                    data$Month == "2021_Dec"), ]

# Store 16 all data:
data_16_allmonths <- data[!(data$Store_Id == "1" | data$Store_Id == "23"), ]


# Store 1 train model data:
data_1 <- data[!(data$Store_Id == "16" | data$Store_Id == "23" | 
                   data$Month == "2021_Nov" | data$Month == "2021_Oct"| 
                    data$Month == "2021_Dec"), ]

# Store 1 all data:
data_1_allmonths <- data[!(data$Store_Id == "16" | data$Store_Id == "23"), ]


```


Store 16: Create general additive models with smoothing terms for month and seasonal sales cycles.
```{r}

# Tire sales: 
model_gam_sales <- gam(Tire_Sales ~ s(Month_count, k = 6, bs = 'cs') + 
                            s(Month_count2, k = 5, bs = 'cs'), 
                            data = data_16)  # model

data_16_allmonths$fit_sales <- predict(model_gam_sales, 
                            data_16_allmonths, 
                            type = "response")  # estimates



# Tire services: 
model_gam_services <- gam(Service_sales ~ s(Month_count, k = 6, bs = 'cs') + 
                            s(Month_count2, k = 5, bs = 'cs'), 
                            data = data_16)  # model

data_16_allmonths$fit_services <- predict(model_gam_services, 
                            data_16_allmonths, 
                            type = "response")  # estimates


```


Store 16: Visualize the models and forecast revenue for each sale type:
```{r}

# Tires Sales

ggplot() +
  
  geom_point(aes(x = data_16$Month, y = data_16$Tire_Sales), # Sales
                  alpha = 0.5, colour = 'red') +
  geom_point(aes(x = data_16_allmonths$Month, y = data_16_allmonths$fit_sales), # Forecast sales
                  alpha = 0.4,colour = "blue") +

  geom_smooth(aes(x = data_16_allmonths$ Month, 
                  y = data_16_allmonths$fit_sales, group = 1), # Visualize the model
                  method = 'gam', lty = 'blank', fill = "blue", 
                  alpha = 0.1, se = TRUE) +

  labs(x = "", y = "Tire Sales", color = "Legend") +

  ylim(0, 20000) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  scale_x_discrete(limits = data_16_allmonths$Month) 



# Service Sales

ggplot() +
  
  geom_point(aes(x = data_16$Month, y = data_16$Service_sales),  # Sales
                  alpha = 0.5, colour = 'red') +
  geom_point(aes(x = data_16_allmonths$Month, y = data_16_allmonths$fit_services), # Forecast sales
                  alpha = 0.4, colour = "blue") +

  geom_smooth(aes(x = data_16_allmonths$Month,
                  y = data_16_allmonths$fit_services, group = 1), 
                  method = 'gam', lty = 'dotted', fill = "blue", 
                  alpha = 0.1, se = FALSE) +

  labs(x = "", y = "Tire Service Sales", color = "Legend") +

  ylim(0, 7500) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  scale_x_discrete(limits = data_16_allmonths$Month) 



```


Store 1: Create general additive models with smoothing terms for month and seasonal sales cycles.
```{r}

# Tire sales:
model_gam_sales <- gam(Tire_Sales ~ s(Month_count, k = 5, bs = 'cc') +
                        s(Month_count2, k = 7, bs = 'cs'), data = data_1)  # model
data_1_allmonths$fit_sales <- predict(model_gam_sales, data_1_allmonths, type = "response")  # estimates


# Tire services:
model_gam_services <- gam(Service_sales ~ s(Month_count, k = 6, bs = 'cs') +
                            s(Month_count2, k = 5, bs = 'cs'), data = data_1)  # model
data_1_allmonths$fit_services <- predict(model_gam_services, data_1_allmonths, type = "response")  # estimates


```


Store 1: Visualize the models and forecast revenue for each sale type:
```{r}

# Tire sales:

ggplot() +
  
  geom_point(aes(x = data_1_allmonths$Month,  # Sales
                    y = data_1_allmonths$Tire_Sales,
                    colour = data_1_allmonths$Sign), alpha = 0.5) +

  geom_smooth(aes(x = data_1_allmonths$Month,  # Forecast estimates
                    y = data_1_allmonths$fit_sales, group = 1), 
                    method = 'gam', lty = 'dotted', fill = "blue", alpha = 0.1) +

  
  ylim(0, 5000) +
  xlab("") +
  theme_gray() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  scale_x_discrete(limits = data_1_allmonths$Month) +
  ylab("Tire Sales")



#  Service sales:

ggplot() +
  
  geom_point(aes(x = data_1_allmonths$Month,   # Sales
                    y = data_1_allmonths$Service_sales), 
                    alpha = 0.5) +
  


  geom_smooth(aes(x = data_1_allmonths$Month,  # Forecast estimates
                    y = data_1_allmonths$fit_services, group = 1), 
                    method = 'gam', lty = 'dotted', fill = "blue", alpha = 0.1) +

  
  ylim(0, 5000) +
  xlab("") +
  theme_gray() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  scale_x_discrete(limits = data_1_allmonths$Month) +
  ylab("Tire Service Sales")



```






